---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# R/`origami`

[![Travis-CI Build Status](https://travis-ci.org/jeremyrcoyle/origami.svg?branch=master)](https://travis-ci.org/jeremyrcoyle/origami)
[![Build status](https://ci.appveyor.com/api/projects/status/i5qwp8cjb4j4x329?svg=true)](https://ci.appveyor.com/project/jeremyrcoyle/origami)
[![Coverage Status](https://img.shields.io/codecov/c/github/jeremyrcoyle/origami/master.svg)](https://codecov.io/github/jeremyrcoyle/origami?branch=master)
[![CRAN](http://www.r-pkg.org/badges/version/origami)](http://www.r-pkg.org/pkg/origami)
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![License: GPL v3](https://img.shields.io/badge/License-GPL%20v3-blue.svg)](http://www.gnu.org/licenses/gpl-3.0)

> High-powered framework for cross-validation: fold your data like it's paper!

__Authors:__ [Jeremy Coyle](https://github.com/jeremyrcoyle) and [Nima
Hejazi](http://nimahejazi.org)

---

## Description

`origami` is an R package that provides a general framework for the application
of cross-validation schemes to particular functions. By allowing arbitrary
lists of results, `origami` accommodates a range of cross-validation applications.

---

## Installation

For standard use, we recommend installing the package from
[CRAN](https://cran.r-project.org/) via

```{r cran-installation, eval = FALSE}
install.packages("survtmle")
```

You can install a stable release of `origami` from GitHub via
[`devtools`](https://www.rstudio.com/products/rpackages/devtools/) with:

```{r gh-installation, eval = FALSE}
devtools::install_github("jeremyrcoyle/origami")
```

---

## Usage

For details on how best to use `origami`, please consult the package vignette
from within R, or browse the vignette
[here](http://nimahejazi.org/origami/articles/generalizedCV.html).

---

## Example

This minimal example shows how to use `origami` to apply cross-validation to the
computation of a simple descriptive statistic using a sample data set. In
particular, we obtain a cross-validated estimate of the mean:

```{r example}
data(mtcars)
head(mtcars)
library(origami)

cvlm <- function(fold) {
  train_data <- training(mtcars)
  valid_data <- validation(mtcars)

  mod <- lm(mpg ~ ., data = train_data)
  preds <- predict(mod, newdata = valid_data)
  list(coef = data.frame(t(coef(mod))), SE = ((preds - valid_data$mpg)^2))
}

folds <- make_folds(mtcars)
results <- cross_validate(cvlm, folds)
mean(results$SE)
```

---

## Issues

If you encounter any bugs or have any specific feature requests, please [file an
issue](https://github.com/jeremyrcoyle/origami/issues).

---

## Citation

After using the `origami` R package, please cite it:

        @article{coyle2017origami,
          doi = {},
          url = {},
          year  = {2017},
          month = {},
          publisher = {The Open Journal},
          volume = {},
          number = {},
          author = {Coyle, Jeremy R and Hejazi, Nima S},
          title = {origami: A Generalized Framework for Cross-Validation in R},
          journal = {The Journal of Open Source Software}
        }

---

## License

&copy; 2017 [Jeremy R. Coyle](https://github.com/jeremyrcoyle)

The contents of this repository are distributed under the GPL-3 license. See
file `LICENSE` for details.
